Turbo fast no compromise gravitational wave parameter estimation

### 2022-11-10 src/tex/ms.tex
update tex
Update intorduction

\section{Gravitational wave parameter estimation}

@@

\subsection{MCMC}

\begin{align}
    p(\theta| d) = \frac{\mathcal{L}(d|\theta)p(\theta)}{p(d)}
\end{align}

@@

\begin{align}
    \left<a|b\right> = 4 Re\Int \frac{a(f)^*b(f)}{\mathcal{S}_n(f)} df
\end{align}

\subsection{Waveform}

@@

\section{FlowMC}

@@

Since our likelihood function is coded in 
\subsection{Gradient-based sampler}
\label{sec:gradient}

\texttt{Jax} allows us to compute the gradient of the likelihood function with
respect to the parameters through automatic differentiation. This enables us to
use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\cite{kw}. The main advantage
of gradient-based sampling methods over gradient-free sampling methods is the
gradient information help decorrelates different parameters. This means
gradient-based sampling methods allow longer jump in a high dimensional space.
This is crucial for high dimensional problems \footnote{Gradient-free methods
often have troubles dealing with problems with $>20$ dimensions.}, since
gradient-free methods would need small step size to have reasonable acceptance
rate in a high dimensional space, which in turn leads to a more correlated
samples hence slower convergence.

\kw{Show some examples of successful application of HMC}

@@

% Long burn in too
Moreover, before we can use the sampling chain to estimate the quantity we care,
the sampler often needs to first find the most probable region in the target
space, which is a common process that is referred to "burn-in" in the
literature. This means one would discard a certain amount of data generated from
the beginning and only use the later part to estimate the quantities of
interest. The burn-in phase of a gradient-based sampler is often as long as the
sampling phase, which means a good portion of computation is not necessary
helpful in estimating the target quantities.

% Crux of normalizing flow
Normalizing flow is a neural-network based technique that aims at learning a
mapping between a simple distribution, such as a Gaussian distribution, to a
complex distribution, often given in the form of data samples from a target
distribution. Once the network is trained, one can evaluate the probability
density of the complex distribution and samples from it very efficiently.
The core of normalizing flow is given by
\begin{align}
    p_x(X) = p_z(Z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|,
\end{align}
where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
distribution, and $f$ is a series of invertible parameterized transform that
connect the two distributions. For a detail discussion of the algorithm, we
refer the readers to \kw{cite}.

@@

@@

% More sample helps training normalizing flow
In this work, the use of accelerators provides two independent perks that
tremendously benefits the parameter estimation process. First, using
accelerators allow us to run many independent chains at the same time, which
benefits the training of the normalizing flow. Since we generate the data we use
to train the normalizing flow on-the-fly, the more independent data we can feed
to the training process, the higher chance the normalizing flow can learn a
reasonable representation of the global landscape of the target distribution. If
we only use a small amount of chains, we are limited to the correlated samples
from each chain, so we have to run more sequential steps to get the same amount
of independent data as compared to running more chains, where former option does
not benefit from parallelization but the latter does. In another word, being
able to use many independent chains help the normalizing flow learn the global
landscape faster in WALL time.

@@

\section{Checks}

@@

@@

\section{Discussion}

>>>>

% Brief description of PE and old generation of code
Parameter estimation (PE) is one of the most commonly perform data analysis
tasks in gravitational wave (GW) data analysis. The central question of PE is to
infer the parameters of the GW source given the strain data. There are a number
of standard tools developed by the community \kw{cite} tha

% Emerging tools from different group
While the

%Basic idea and unique advantage of our tool

%Structure of paper

The rest of the paper is structured as follows: We review the basic of PE in
section \ref{sec: PE}. We provide a quick summary of the sampling framework
built on top of \texttt{flowMC} in section \ref{sec: flowMC}. We present
benchmarking result in section \ref{sec: Result}. Finally, we discuss the
implications of this work and directions for future development in section
\ref{sec: Discussion}.

\section{Gravitational wave parameter estimation}
\label{sec: PE}

@@

\subsection{Likelihood function}
\label{sec:likelihood}

\begin{align}
    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)}
\end{align}

@@

\begin{align}
    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df
\end{align}

% I think we should put a short description of MALA here
Since our likelihood function is coded in 
\subsection{MCMC with Gradient-based sampler}
\label{sec:gradient}

\texttt{Jax} allows us to compute the gradient of the likelihood function with
respect to the parameters through automatic differentiation. This enables us to
use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\kw{cite}. The main advantage
of gradient-based sampling methods over gradient-free sampling methods is the
gradient information help decorrelates different parameters. This means
gradient-based sampling methods allow longer jump in a high dimensional space.
This is crucial for high dimensional problems \footnote{Gradient-free methods
often have troubles dealing with problems with $>20$ dimensions.}, since
gradient-free methods would need small step size to have reasonable acceptance
rate in a high dimensional space, which in turn leads to a more correlated
samples hence slower convergence.

@@

\begin{align}
r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots
\end{align}

\begin{align}
    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b) \nonumber \\
    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)]
\end{align}

\section{\texttt{flowMC}}
\label{sec: flowMC}

@@

@@

% Long burn-in too
Moreover, before we can use the sampling chain to estimate the quantity we care,
the sampler often needs to first find the most probable region in the target
space, which is a common process that is referred to "burn-in" in the
literature. This means one would discard a certain amount of data generated from
the beginning and only use the later part to estimate the quantities of
interest. The burn-in phase of a gradient-based sampler is often as long as the
sampling phase, which means a good portion of computation is not necessary
helpful in estimating the target quantities.

% Crux of normalizing flow
Normalizing flow is a neural-network based technique that aims at learning a
mapping between a simple distribution, such as a Gaussian distribution, to a
complex distribution, often given in the form of data samples from a target
distribution. Once the network is trained, one can evaluate the probability
density of the complex distribution and samples from it very efficiently.
The core of normalizing flow is given by
\begin{align}
    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|,
\end{align}
where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
distribution, and $f$ is an invertible parameterized transform that
connect the two distributions. For a detail discussion of the algorithm, we
refer the readers to \kw{cite}.

@@

% Write down how you use it as proposal

\begin{align}
    acc(x,x_0) = \rm{min} \left[ 1, \frac{\hat{\rho}(x_0)\rho_*(x)}{\hat{\rho}(x)\rho_*(x_0)}\right],
\end{align}
where $x = f(z), z \sim p_z(Z)$ is the proposed jump location and $x_0$ is the current position.
We can see the flow distribution is the target distribution when the accepting probability is 1.

@@

% More sample helps training normalizing flow
In this work, the use of accelerators provides two independent perks that
tremendously benefits the parameter estimation process. First, using
accelerators allow us to run many independent chains at the same time, which
benefits the training of the normalizing flow. Since we generate the data we use
to train the normalizing flow on-the-fly, the more independent data we can feed
to the training process, the higher chance the normalizing flow can learn a
reasonable representation of the global landscape of the target distribution. If
we only use a small amount of chains, we are limited to the correlated samples
from each chain, so we have to run more sequential steps to get the same amount
of independent data as compared to running more chains, where former option does
not benefit from parallelization but the latter does. In another word, being
able to use many independent chains help the normalizing flow learn the global
landscape faster in Wall time.

@@

\section{Result}
\label{sec: Result}

In this section, we provide a number of sanity checks as well as a few real life applications

@@

\subsection{Injection-recovery test}

\subsection{Real event parameter estimation}

@@

\section{Discussion}
\label{sec: Discussion}
